Distributed Nonlinear Control Design using Separable Control Contraction Metrics

### 2017-06-01 R0V0.tex
Merge Ian's changes in the introduction

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
In this paper, large-scale input-affine nonlinear systems are considered. The problem is the design of distributed controllers for this class of network systems. Although control-Lyapunov functions are the natural candidates to design feedback laws, the computation of these functions is a non-convex optimization problem. Moreover, it may require to verify the positiveness of a function with, at least, $n^2$ elements. In this paper, control-contraction metrics (CCMs) are employed to design controllers with prescribed distributed structures. By using a separable structure, the number of verifications needed can be made smaller, as the search of CCMs is formulated as a distributed semidefinite program. An example illustrates the approach for a network system.
\end{abstract}

@@

\mypara Over the years, improvements on distance communication and cheaper computation platforms have allowed the development of large-scale interconnected systems such as smart grids, sensor networks \cite{Pajic2011}, smart manufacturing plants, traffic networks \cite{CanudasdeWitMorbidiLeonOjedaEtAl2015} and pipe networks \cite{Persis2011a}. One of the main characteristic of these systems is the capacity of being autonomous. Most of current control algorithms are based on the assumption that the system will operate ``near'' to a prescribed set point. Consequently, the system is designed to have fixed behavior. This assumption imposes severe limitations for systems, as large excursions are not allowed, in general. For instance, large power variations in smart grids (e.g., when renewable sources of energy are taken into account) are not allowed and manufacturing plants tend to synthesize only a fixed range of goods. To render these systems more flexible, i.e., allowing large excursions a controller algorithm that takes into account the behavior far from the set point is needed.

\mypara In terms of control theory, the difference between a system operating in a set point and operating in a much larger region lies in the type of equation that describes the system behavior. In the former, the system is described by linear equation while, in the latter, by a nonlinear one. In general, controller design methods for nonlinear systems require the knowledge of a \emph{Lyapunov function}. The search of Lyapunov functions for this class of systems is a non-trivial problem by itself, as no general methodology exists. Another drawback of Lyapunov functions is the fact that they are defined with respect to a particular set point. In case of large excursions, it is more appropriate to define a function depending on the distance between points. Even if a Lyapunov function is known, no general methodology is known to design distributed feedback laws for nonlinear systems. Thus, existing approaches are, in general, \emph{ad-hoc}. When the design of distributed controllers is considered, the situation is even worse as the problem is known to be, at least, NP-hard \cite{BlondelTsitsiklis1997,Tanaka2011}.

\mypara  An alternative method, called control contraction method (CCM), provides an algorithm in which nonlinear feedback laws are designed based on solving a convex optimization problem  \cite{Manchester2014a}. The main challenge with this methodology is that, when the system under consideration has a large number of states, it may require a large memory amount. This is due to the fact that the stability verification of a linear system with $n$ states may require a function with, at least, $n^2$ terms (see \cite{Rantzer2015} for the case of linear systems and Lyapunov functions).

\mypara One of the key elements employed to design these controllers is the concept of a Riemannian metric (see the definition below) which is employed, similarly to Lyapunov functions, to measure the exponential convergence of pairs of solutions to each other. Employing CCMs that can be decomposed into the sum of smaller elements, the design of a distributed feedback law allowing exponential convergence of solutions is obtained in two steps.

\mypara The first step consists of an off-line computation. The original system is extended with its \emph{differential / variational} formulation. The exponential convergence problem is reformulated in terms of stability of the origin for the extended system. This stability problem is written in terms of a convex optimization problem under the prescribed structural constraint. The solution consists of a pair of matrices: the Riemannian metric and the \emph{differential / variational} feedback law. The second step consists of an on-line optimization. The obtained \emph{differential / variational} controller is integrated along geodesics (see definition below).  In contrast to feedback linearization which depends on integrability, the obtained controller depends only on the stabilizability property of the system. This stability notion is related to contraction theory (an incremental stability property \cite{Angeli2002,Lohmiller1998,Rueffer2013,Sontag2010,Forni2014}) which can be traced back to the work \cite{Lewis1949}.

\mypara At this point, the three contributions of this work can be highlighted. First, using the methodology for synthesis of CCMs introduced in \cite{Manchester2014a} and imposing a (block-)diagonal structural constraint, a distributed feedback law is obtained. Second, because the CCM has a (block-)diagonal structure, the computation of the online integration phase, which corresponds to an optimization problem, is also distributed. Third, the structure imposed on the CCM allows to reduce the number of equations needed to  verify stability. Note that the proposed method provides an explicit algorithm to design the controller.

\mypara Other works related to analysis under structural constraints to achieve exponential convergence of solutions do exist in the literature. For contraction analysis, see \cite{Aminzare2014a} for analysis according to different network topologies and \cite{Andrieu2016} for transversal contraction analysis. This paper contrast with these works mainly in two points: 1) A  methodology for the design of structured controllers is proposed; 2) Due the formulation of the stabilization problem as a convex optimization problem, an explicit algorithm of the distributed controller is provided. A particular case of this paper has also been published in \cite{SteinShiromotoManchester2016}. In contrast to this paper, more general conditions to design the distributed controller (cf. Corollary~\ref{cor:main result}) and a methodology to distribute the computation of the feedback law are given.

\paragraph{Overview} The motivation and the problem formulation are provided in Section \ref{sec:Problem Formulation and Motivation}. The background needed for this paper is stated in Section \ref{sec:background}. Illustrations of the proposed approach are provided in Section \ref{sec:Illustration}. Section \ref{sec:Conclusion} collects final remarks.

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
This paper gives convex conditions for synthesis of a distributed  control system control for a class of large-scale networked nonlinear dynamic systems.  It is shown that the technique of control contraction metrics (CCMs) can be extended to this problem by utilizing {\em separable} metric structures, resulting in controllers that only depend on information from local sensors and communications from immediate neighbours. The conditions given are pointwise linear matrix inequalities, and are necessary and sufficient for linear positive systems. Distributed synthesis methods based on chordal graphs are also proposed. A simple example demonstrates feasibility for network  of nonlinear dynamic systems with up to 512 nodes.
\end{abstract}

@@

\mypara In recent years, rapid advances in communication and  computation technology have enabled the development of large-scale engineered systems such as smart grids \cite{hill_smart_2012}, sensor networks \cite{Pajic2011}, smart manufacturing plants \cite{wang_implementing_2016}, traffic networks \cite{CanudasdeWitMorbidiLeonOjedaEtAl2015} and pipe networks \cite{Persis2011a}. Despite these advances, the design of feedback controllers for such large systems remains challenging.

\mypara When it is assumed that a system  has linear dynamics and that all sensor information can be collected in a single location for control computation, well-developed synthesis methods such as LQG and $H^\infty$ can be applied \cite{anderson1990optimal, dullerud2000course}. However, emerging applications motivate going beyond these assumptions:
\begin{enumerate}[i)]
	\item For geographically distributed systems with thousands or millions of nodes, such as transportation and power networks, it is not practical to collect all sensor information in one location for control. In this case there is a need for {\em distributed} methods that rely only on local information or information communicated from nearby nodes.
 \item Most real systems exhibit {\em nonlinear} dynamics. The use of linear dynamic models is usually based on the premise that the system always operates near a prescribed set-point. When large excursions in operating conditions are expected, e.g. due to changing production demands in a flexible manufacturing system, or recovery from a fault in a smart electrical grid, one must take into account the system nonlinearity.
\end{enumerate}

%Most of current control algorithms are based on the assumption that the system will operate ``near'' to a prescribed set point. Consequently, the system is designed to have fixed behavior. This assumption imposes severe limitations for systems, as large excursions are not allowed, in general. For instance, large power variations in smart grids (e.g., when renewable sources of energy are taken into account) are not allowed and manufacturing plants tend to synthesize only a fixed range of goods. To render these systems more flexible, i.e., allowing large excursions a controller algorithm that takes into account the behavior far from the set point is needed.

\mypara Decentralised and distributed control are long-standing problems in control theory, with important early work surveyed in \cite{sandell_survey_1978}. Terminology is not completely uniform in the literature, but
in this paper we take ``decentralised'' to mean that at each node there is a controller that uses {\em only} local information, and  ``distributed'' to mean that {\em some} communication is allowed between nearby nodes. In either case, there is a desire to impose a {\em structure} on information flow, and this turns out to be the main source of difficulty.

\mypara Even for linear systems, it has long been known that apparently simple problems with decentralized information flow can be surprisingly challenging \cite{witsenhausen_counterexample_1968}. A recent breakthrough was the discovery that a property called {\em quadratic invariance} characterises a convex subclass of problems in which, roughly speaking, communication is faster than the propagation of dynamical effects \cite{rotkowitz_characterization_2006}.
%
%\begin{align}
%M(A+BK)+(A+BK)'M<0\\
%\Updownarrow\\
%AW+WA'+BY+Y'B'<0	
%\end{align}

\mypara More directly related to this paper is a large body of work on linear matrix inequality (LMI) methods. For linear state feedback, information flow can be encoded by a sparsity structure on the feedback gain matrix, however in general the problem of finding such a gain which is stabilizing is NP-Hard \cite{BlondelTsitsiklis1997}. It has been recognized by many authors that if the search is restricted to {\em diagonal} (or {\em block diagonal}) Lyapunov matrices, then the problem is convex (see, e.g., \cite{zecevic_control_2010, Tanaka2011, Rantzer2015} and references therein). The main benefit is that sparsity structure in the gain matrix is preserved under the standard change of variables for LMI-based design.

\mypara In general, restricting the set of Lyapunov functions is conservative: it produces sufficient conditions for stabilizability, but not necessary. However, for the important sub-class of systems for which internal states are always  non-negative, known as {\em positive systems}, existence of a diagonal Lyapunov function is actually necessary and sufficient
(see, e.g., \cite{berman_nonnegative_1994} and references therein). This result has been extended to $H^\infty$ design \cite{Tanaka2011}, robust stability \cite{colombino_convex_2016}, and scalable algorithms for control design \cite{Rantzer2015} and identification \cite{umenberger_scalable_2016} of networked positive systems.

\mypara Design of controllers for nonlinear systems has also been a major topic of research for many years, see e.g. \cite{Slotine1991,Krstic1995, Sontag1998, Khalil:2001} for established approaches.
Most methods require (at least implicitly) the construction of a \emph{control Lyapunov function}. While for certain structured systems, constructive methods such as backstepping and energy-based control can be used \cite{Slotine1991, Krstic1995}, no general methodology exists. Indeed, the set of control Lyapunov functions can be non-convex and disconnected \cite{Rantzer:2001}.

Separable Lyapunov functions
\cite{Dirr2015} and references therein. \cite{liu_nonlinear_2014}.

\mypara Another drawback of Lyapunov functions is the fact that they are defined with respect to a particular set point or trajectory, which must be known {\em a priori}. In many cases, e.g. robotics or flexible manufacturing, it is more appropriate to define a function depending on the distance {\em between} pairs of points. Tools such as contraction metrics \cite{Lohmiller1998} and incremental Lyapunov functions \cite{Angeli2002} provide such a capability for stability analysis, but the problem of going from analysis to synthesis remains.

\mypara  An alternative approach is to construct a control contraction metric (CCM), introduced in \cite{Manchester2014a, manchester_control_2017}. The main advantages this method offers over the Lyapunov approach are that the synthesis conditions are convex, and that provides a stabilizing controller for {\em all} solutions, not just a single set point. It was shown in \cite{manchester_control_2017} that the CCM conditions are necessary and sufficient for feedback-linearizable nonlinear systems.

%mypara One of the key elements employed to design these controllers is the concept of a Riemannian metric (see the definition below) which is employed, similarly to Lyapunov functions, to measure the exponential convergence of pairs of solutions to each other. Employing CCMs that can be decomposed into the sum of smaller elements, the design of a distributed feedback law allowing exponential convergence of solutions is obtained in two steps.

%\mypara The first step consists of an off-line computation. The original system is extended with its \emph{differential / variational} formulation. The exponential convergence problem is reformulated in terms of stability of the origin for the extended system. This stability problem is written in terms of a convex optimization problem under the prescribed structural constraint. The solution consists of a pair of matrices: the Riemannian metric and the \emph{differential / variational} feedback law. The second step consists of an on-line optimization. The obtained \emph{differential / variational} controller is integrated along geodesics (see definition below).  In contrast to feedback linearization which depends on integrability, the obtained controller depends only on the stabilizability property of the system. This stability notion is related to contraction theory (an incremental stability property \cite{Angeli2002,Lohmiller1998,Rueffer2013,Sontag2010,Forni2014}) which can be traced back to the work \cite{Lewis1949}.

Contraction analysis of networked systems
\cite{Russo2013}
\cite{Aminzare2014a} 
\cite{manchester_existence_2018}

\mypara At this point, the three contributions of this work can be highlighted. First, using the methodology for synthesis of CCMs introduced in \cite{manchester_control_2017} and imposing a (block-)diagonal structural constraint, a distributed feedback law is obtained. Second, because the CCM has a (block-)diagonal structure, the computation of the online integration phase, which corresponds to an optimization problem, is also distributed. Third, the structure imposed on the CCM allows to reduce the number of equations needed to  verify stability. Note that the proposed method provides an explicit algorithm to design the controller.

A particular case of this paper has also been published in \cite{SteinShiromotoManchester2016}

\paragraph{Overview} The motivation and the problem formulation are provided in Section \ref{sec:Problem Formulation and Motivation}, with technical background in Section \ref{sec:background}. Sections \ref{sec:Design of Decentralized Controllers} and \ref{sec:distributed_design} present the proposed method and main theoretical results. Illustrations of the proposed approach are provided in Section \ref{sec:Illustration}. Section \ref{sec:Conclusion} collects final remarks.

\section{Preliminaries and Problem Formulation}
